PANCHANGE_data_preprocessing

### 2020-09-09 PANCHANGE_data_preprocessing2020-07-15.Rmd

>>>>

# Neurological symptoms

## ALS (alsfrs)

### GLAD
Read in data
```{r glad als cols}
if(GLAD == TRUE){
glad.als.raw <- readRDS(file = paste0(data.raw_path, "/glad/alsfrs_coping_glad.rds"))
dim(glad.als.raw)
colnames(glad.als.raw)
}
```

Select variables
```{r glad als cleaning}
if(GLAD == TRUE){

glad.als.raw.id <- glad.als.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
         Sample, # Sample
         ID, # ID
         alsfrs.limb_weakness_noticed_past,
         alsfrs.normal,
         alsfrs.slow_or_sloppy_all_words_are_legible,
         alsfrs.not_all_words_are_legible,
         alsfrs.able_to_grip_pen,
         alsfrs.unable_to_grip_pen,
         alsfrs.have_you_noticed_any_difficulty_walking,
         alsfrs.do_you_have_any_difficulty_climbing_stairs
         ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(glad.als.raw.id)
# Inspect colnames
colnames(glad.als.raw.id)
#Differences
dim(glad.als.raw)[1]-dim(glad.als.raw.id)[1]
}
```

```{r glad copy als}
if(GLAD == TRUE){
als.raw.id <- glad.als.raw.id
}
```

### EDGI
Read in data
```{r edgi als cols}
if(EDGI == TRUE){
edgi.als.raw <- readRDS(file = paste0(data.raw_path, "/edgi/alsfrs_coping_edgi.rds"))
dim(edgi.als.raw)
colnames(edgi.als.raw)
}
```

Select variables
```{r edgi als cleaning}
if(EDGI == TRUE){

edgi.als.raw.id <- edgi.als.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
         Sample, # Sample
         ID, # ID
         alsfrs.limb_weakness_noticed_past,
         alsfrs.normal,
         alsfrs.slow_or_sloppy_all_words_are_legible,
         alsfrs.not_all_words_are_legible,
         alsfrs.able_to_grip_pen,
         alsfrs.unable_to_grip_pen,
         alsfrs.have_you_noticed_any_difficulty_walking,
         alsfrs.do_you_have_any_difficulty_climbing_stairs
         ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(edgi.als.raw.id)
# Inspect colnames
colnames(edgi.als.raw.id)
#Differences
dim(edgi.als.raw)[1]-dim(edgi.als.raw.id)[1]
}
```

```{r edgi copy als}
if(EDGI == TRUE){
als.raw.id <- edgi.als.raw.id
}
```

### NBR
Read in data
```{r nbr als cols}
if(NBR == TRUE){
nbr.als.raw <- readRDS(file = paste0(data.raw_path, "/nbr/alsfrs_coping_nbr.rds"))
dim(nbr.als.raw)
colnames(nbr.als.raw)
}
```

Select variables
```{r nbr als cleaning}
if(NBR == TRUE){

nbr.als.raw.id <- nbr.als.raw %>%
  drop_na(subjectid) %>% # Drop NAs
  distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(subjectid, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
         Sample, # Sample
         ID, # ID
         alsfrs.limb_weakness_noticed_past,
         alsfrs.normal,
         alsfrs.slow_or_sloppy_all_words_are_legible,
         alsfrs.not_all_words_are_legible,
         alsfrs.able_to_grip_pen,
         alsfrs.unable_to_grip_pen,
         alsfrs.have_you_noticed_any_difficulty_walking,
         alsfrs.do_you_have_any_difficulty_climbing_stairs
         ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(nbr.als.raw.id)
# Inspect colnames
colnames(nbr.als.raw.id)
#Differences
dim(nbr.als.raw)[1]-dim(nbr.als.raw.id)[1]
}
```

```{r nbr copy als}
if(NBR == TRUE){
als.raw.id <- nbr.als.raw.id
}
```

### RAMP
Read in data
```{r ramp als cols}
if(RAMP == TRUE){
ramp.als.raw <- readRDS(file = paste0(data.raw_path, "/ramp/alsfrs_ramp.rds"))
dim(ramp.als.raw)
colnames(ramp.als.raw)
}
```

Select variables
```{r ramp als cleaning}
if(RAMP == TRUE){

ramp.als.raw.id <- ramp.als.raw %>%
  drop_na(`Login ID`) %>% # Drop NAs
  distinct(`Login ID`, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(`Login ID`, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
         Sample, # Sample
         ID, # ID
         alsfrs.limb_weakness_noticed_past,
         alsfrs.normal,
         alsfrs.slow_or_sloppy_all_words_are_legible,
         alsfrs.not_all_words_are_legible,
         alsfrs.able_to_grip_pen,
         alsfrs.unable_to_grip_pen,
         alsfrs.have_you_noticed_any_difficulty_walking,
         alsfrs.do_you_have_any_difficulty_climbing_stairs
         ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(ramp.als.raw.id)
# Inspect colnames
colnames(ramp.als.raw.id)
#Differences
dim(ramp.als.raw)[1]-dim(ramp.als.raw.id)[1]
}
```

```{r ramp copy als}
if(RAMP == TRUE){
als.raw.id <- ramp.als.raw.id
}
```

## Epsworth sleep scale (ess)

### GLAD
Read in data
```{r glad ess cols}
if(GLAD == TRUE){
glad.ess.raw <- readRDS(file = paste0(data.raw_path, "/glad/ess_coping_glad.rds"))
dim(glad.ess.raw)
colnames(glad.ess.raw)
}
```

Select variables
```{r glad ess cleaning}
if(GLAD == TRUE){

glad.ess.raw.id <- glad.ess.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
         Sample, # Sample
         ID, # ID
         ess.felt_excessively_sleepy_past= ess.day_past_month_felt,
         ess.sitting_and_reading,
         ess.watching_tv,
         ess.sitting_inactive_in_a_public_place,
         ess.break_passenger_hour_car,
         ess.afternoon_rest_circumstances_lying,
         ess.sitting_and_talking_to_someone,
         ess.sitting_quietly_after_a_lunch_without_alcohol,
         ess.car_minutes_stopped_traffic,
         ess.insomnia,
         ess.narcolepsy,
         ess.rem_sleep_behaviour_disorder,
         ess.restless_legs_syndrome,
         ess.sleep_apnea,
         ess.sleep_terrors,
         ess.sleepwalking,
         ess.other,
         ess.none   
         ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(glad.ess.raw.id)
# Inspect colnames
colnames(glad.ess.raw.id)
#Differences
dim(glad.ess.raw)[1]-dim(glad.ess.raw.id)[1]
}
```

```{r glad copy ess}
if(GLAD == TRUE){
ess.raw.id <- glad.ess.raw.id
}
```

### EDGI
Read in data
```{r edgi ess cols}
if(EDGI == TRUE){
edgi.ess.raw <- readRDS(file = paste0(data.raw_path, "/edgi/ess_coping_edgi.rds"))
dim(edgi.ess.raw)
colnames(edgi.ess.raw)
}
```

Select variables
```{r edgi ess cleaning}
if(EDGI == TRUE){

edgi.ess.raw.id <- edgi.ess.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
         Sample, # Sample
         ID, # ID
         ess.felt_excessively_sleepy_past,
         ess.sitting_and_reading,
         ess.watching_tv,
         ess.sitting_inactive_in_a_public_place,
         ess.break_passenger_hour_car,
         ess.afternoon_rest_circumstances_lying,
         ess.sitting_and_talking_to_someone,
         ess.sitting_quietly_after_a_lunch_without_alcohol,
         ess.car_minutes_stopped_traffic,
         ess.insomnia,
         ess.narcolepsy,
         ess.rem_sleep_behaviour_disorder,
         ess.restless_legs_syndrome,
         ess.sleep_apnea,
         ess.sleep_terrors,
         ess.sleepwalking,
         ess.other,
         ess.none  
         ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(edgi.ess.raw.id)
# Inspect colnames
colnames(edgi.ess.raw.id)
#Differences
dim(edgi.ess.raw)[1]-dim(edgi.ess.raw.id)[1]
}
```

```{r edgi copy ess}
if(EDGI == TRUE){
ess.raw.id <- edgi.ess.raw.id
}
```

### NBR
Read in data
```{r nbr ess cols}
if(NBR == TRUE){
nbr.ess.raw <- readRDS(file = paste0(data.raw_path, "/nbr/ess_coping_nbr.rds"))
dim(nbr.ess.raw)
colnames(nbr.ess.raw)
}
```

Select variables
```{r nbr ess cleaning}
if(NBR == TRUE){

nbr.ess.raw.id <- nbr.ess.raw %>%
  drop_na(subjectid) %>% # Drop NAs
  distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(subjectid, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
         Sample, # Sample
         ID, # ID
         ess.felt_excessively_sleepy_past,
         ess.sitting_and_reading,
         ess.watching_tv,
         ess.sitting_inactive_in_a_public_place,
         ess.break_passenger_hour_car,
         ess.afternoon_rest_circumstances_lying,
         ess.sitting_and_talking_to_someone,
         ess.sitting_quietly_after_a_lunch_without_alcohol,
         ess.car_minutes_stopped_traffic,
         ess.insomnia,
         ess.narcolepsy,
         ess.rem_sleep_behaviour_disorder,
         ess.restless_legs_syndrome,
         ess.sleep_apnea,
         ess.sleep_terrors,
         ess.sleepwalking,
         ess.other,
         ess.none  
         ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(nbr.ess.raw.id)
# Inspect colnames
colnames(nbr.ess.raw.id)
#Differences
dim(nbr.ess.raw)[1]-dim(nbr.ess.raw.id)[1]
}
```

```{r nbr copy ess}
if(NBR == TRUE){
ess.raw.id <- nbr.ess.raw.id
}
```

### RAMP
Read in data
```{r ramp ess cols}
if(RAMP == TRUE){
ramp.ess.raw <- readRDS(file = paste0(data.raw_path, "/ramp/ess_ramp.rds"))
dim(ramp.ess.raw)
colnames(ramp.ess.raw)
}
```

Select variables
```{r ramp ess cleaning}
if(RAMP == TRUE){

ramp.ess.raw.id <- ramp.ess.raw %>%
  drop_na(`Login ID`) %>% # Drop NAs
  distinct(`Login ID`, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(`Login ID`, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
         Sample, # Sample
         ID, # ID
         ess.felt_excessively_sleepy_past,
         ess.sitting_and_reading,
         ess.watching_tv,
         ess.sitting_inactive_in_a_public_place,
         ess.break_passenger_hour_car,
         ess.afternoon_rest_circumstances_lying,
         ess.sitting_and_talking_to_someone,
         ess.sitting_quietly_after_a_lunch_without_alcohol,
         ess.car_minutes_stopped_traffic,
         ess.insomnia,
         ess.narcolepsy,
         ess.rem_sleep_behaviour_disorder,
         ess.restless_legs_syndrome,
         ess.sleep_apnea,
         ess.sleep_terrors,
         ess.sleepwalking,
         ess.other,
         ess.none  
         ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(ramp.ess.raw.id)
# Inspect colnames
colnames(ramp.ess.raw.id)
#Differences
dim(ramp.ess.raw)[1]-dim(ramp.ess.raw.id)[1]
}
```

```{r ramp copy ess}
if(RAMP == TRUE){
ess.raw.id <- ramp.ess.raw.id
}
```

## Migraines (midas)

### GLAD
Read in data
```{r glad midas cols}
if(GLAD == TRUE){
glad.midas.raw <- readRDS(file = paste0(data.raw_path, "/glad/midas_coping_glad.rds"))
dim(glad.midas.raw)
colnames(glad.midas.raw)
}
```

Select variables
```{r glad midas cleaning}
if(GLAD == TRUE){

glad.midas.raw.id <- glad.midas.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
         Sample, # Sample
         ID, # ID
         midas.headaches_suffered_past_month,
         midas.days_past_month_miss_work_school = midas.days_past_month,
         midas.days_past_month_work_school_half_productivity = midas.days_past_month.1,
         midas.days_past_month_miss_housework = midas.days_past_month.2,
         midas.days_past_month_housework_half_productivity = midas.days_past_month.3,
         midas.days_past_month_miss_social_activities = midas.days_past_month.4
         ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(glad.midas.raw.id)
# Inspect colnames
colnames(glad.midas.raw.id)
#Differences
dim(glad.midas.raw)[1]-dim(glad.midas.raw.id)[1]
}
```

```{r glad copy midas}
if(GLAD == TRUE){
midas.raw.id <- glad.midas.raw.id
}
```

### EDGI
Read in data
```{r edgi midas cols}
if(EDGI == TRUE){
edgi.midas.raw <- readRDS(file = paste0(data.raw_path, "/edgi/midas_coping_edgi.rds"))
dim(edgi.midas.raw)
colnames(edgi.midas.raw)
}
```

Select variables
```{r edgi midas cleaning}
if(EDGI == TRUE){

edgi.midas.raw.id <- edgi.midas.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
         Sample, # Sample
         ID, # ID
         midas.headaches_suffered_past_month,
         midas.days_past_month_miss_work_school = midas.past_month_days,
         midas.days_past_month_work_school_half_productivity = midas.past_month_days.1,
         midas.days_past_month_miss_housework = midas.past_month_days.2,
         midas.days_past_month_housework_half_productivity = midas.past_month_days.3,
         midas.days_past_month_miss_social_activities = midas.past_month_days.4  
         ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(edgi.midas.raw.id)
# Inspect colnames
colnames(edgi.midas.raw.id)
#Differences
dim(edgi.midas.raw)[1]-dim(edgi.midas.raw.id)[1]
}
```

```{r edgi copy midas}
if(EDGI == TRUE){
midas.raw.id <- edgi.midas.raw.id
}
```

### NBR
Read in data
```{r nbr midas cols}
if(NBR == TRUE){
nbr.midas.raw <- readRDS(file = paste0(data.raw_path, "/nbr/midas_coping_nbr.rds"))
dim(nbr.midas.raw)
colnames(nbr.midas.raw)
}
```

Select variables
```{r nbr midas cleaning}
if(NBR == TRUE){

nbr.midas.raw.id <- nbr.midas.raw %>%
  drop_na(subjectid) %>% # Drop NAs
  distinct(subjectid, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(subjectid, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
         Sample, # Sample
         ID, # ID
         midas.day_past_month_felt,
         midas.sitting_and_reading,
         midas.watching_tv,
         midas.sitting_inactive_in_a_public_place,
         midas.break_passenger_hour_car,
         midas.afternoon_rest_circumstances_lying,
         midas.sitting_and_talking_to_someone,
         midas.sitting_quietly_after_a_lunch_without_alcohol,
         midas.car_minutes_stopped_traffic,
         midas.insomnia,
         midas.narcolepsy,
         midas.rem_sleep_behaviour_disorder,
         midas.restlmidas_legs_syndrome,
         midas.sleep_apnea,
         midas.sleep_terrors,
         midas.sleepwalking,
         midas.other,
         midas.none  
         ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(nbr.midas.raw.id)
# Inspect colnames
colnames(nbr.midas.raw.id)
#Differences
dim(nbr.midas.raw)[1]-dim(nbr.midas.raw.id)[1]
}
```

```{r nbr copy midas}
if(NBR == TRUE){
midas.raw.id <- nbr.midas.raw.id
}
```

### RAMP
Read in data
```{r ramp midas cols}
if(RAMP == TRUE){
ramp.midas.raw <- readRDS(file = paste0(data.raw_path, "/ramp/midas_ramp.rds"))
dim(ramp.midas.raw)
colnames(ramp.midas.raw)
}
```

Select variables
```{r ramp midas cleaning}
if(RAMP == TRUE){

ramp.midas.raw.id <- ramp.midas.raw %>%
  drop_na(`Login ID`) %>% # Drop NAs
  distinct(`Login ID`, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(`Login ID`, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
         Sample, # Sample
         ID, # ID
         midas.day_past_month_felt,
         midas.sitting_and_reading,
         midas.watching_tv,
         midas.sitting_inactive_in_a_public_place,
         midas.break_passenger_hour_car,
         midas.afternoon_rest_circumstances_lying,
         midas.sitting_and_talking_to_someone,
         midas.sitting_quietly_after_a_lunch_without_alcohol,
         midas.car_minutes_stopped_traffic,
         midas.insomnia,
         midas.narcolepsy,
         midas.rem_sleep_behaviour_disorder,
         midas.restlmidas_legs_syndrome,
         midas.sleep_apnea,
         midas.sleep_terrors,
         midas.sleepwalking,
         midas.other,
         midas.none  
         ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(ramp.midas.raw.id)
# Inspect colnames
colnames(ramp.midas.raw.id)
#Differences
dim(ramp.midas.raw)[1]-dim(ramp.midas.raw.id)[1]
}
```

```{r ramp copy midas}
if(RAMP == TRUE){
midas.raw.id <- ramp.midas.raw.id
}
```

## Loss of taste/smell (nmss)

### GLAD
Read in data
```{r glad nmss cols}
if(GLAD == TRUE){
glad.nmss.raw <- readRDS(file = paste0(data.raw_path, "/glad/nmss_coping_glad.rds"))
dim(glad.nmss.raw)
colnames(glad.nmss.raw)
}
```

Select variables
```{r glad nmss cleaning}
if(GLAD == TRUE){

glad.nmss.raw.id <- glad.nmss.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID)) %>%
  select(
         Sample, # Sample
         ID, # ID
         nmss.ability_taste_food_change,
         nmss.severe_taste_smell_loss
         ) %>%
  add_numeric(., exclude = exclude_cols) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(glad.nmss.raw.id)
# Inspect colnames
colnames(glad.nmss.raw.id)
#Differences
dim(glad.nmss.raw)[1]-dim(glad.nmss.raw.id)[1]
}
```

```{r glad copy nmss}
if(GLAD == TRUE){
nmss.raw.id <- glad.nmss.raw.id
}
```

### EDGI
Read in data
```{r edgi nmss cols}
if(EDGI == TRUE){
edgi.nmss.raw <- readRDS(file = paste0(data.raw_path, "/edgi/nmss_coping_edgi.rds"))
dim(edgi.nmss.raw)
colnames(edgi.nmss.raw)
}
```

Select variables
```{r edgi nmss cleaning}
if(EDGI == TRUE){
